Consistency and Fault Tolerance Considerations for the Next Iteration of the DOE Fast Forward Storage and IO Project

### 2014-03-16 paper.tex
version sent to potential team mates.

\documentclass[conference]{sig-alt-gov2}
\pdfpagewidth=8.5in
\pdfpageheight=11in
%\usepackage{subfig}
\usepackage{subfigure}
\usepackage[pdftex]{}
\usepackage{todonotes}
\usepackage{listings}
\lstset {% general command to set parameter(s)
         language=C,
     basicstyle=\footnotesize,               % print whole listing small
%     keywordstyle=\color{black}\bfseries, % underlined bold black keywords
%     identifierstyle =\color{black},  % nothing happens
%     commentstyle=\color{black}\emph, % white comments
     %stringstyle=\ttfamily,          % typewriter type for strings
%     stringstyle=\color{black},       % typewriter type for strings
         tabsize=4,
         showtabs=false,
     showstringspaces=false}
%can't figure this one out for particles bandwidth
%\usepackage[caption,label]{subfig}

@@

@@

\conferenceinfo{MSST'14,} {November 17, 2013, Denver, Colorado, USA.}
\CopyrightYear{2014}
%\crdata{978-1-4503-1103-8/11/11}
\clubpenalty=10000
\widowpenalty = 10000

@@

\numberofauthors{3}
\author{
\alignauthor Jay Lofstead\\
       \affaddr{Sandia National Laboratories}\\
       \email{gflofst@sandia.gov}
\alignauthor Ivo Jimenez\\
       \affaddr{University of California, Santa Cruz}\\
       \email{ivo@cs.ucsc.edu}
\alignauthor Carlos Maltzahn\\
       \affaddr{University of California, Santa Cruz}\\
       \email{carlosm@soe.ucsc.edu}
%\and  % use '\and' if you need 'another row' of author names
}
\maketitle

@@

While these and other their contributions are valuable, similar efforts made
elsewhere may offer attractive alternatives or differing semantics that would
yield a more feature rich environment with little to no additional overhead.
For example, the Doubly Distributed Transactions (\DDTns) protocol offers an
alternative approach for incorporating transactional semantics into the data
path. The PreDatA project examined how to get the best throughput for data
operators can offer some insights for refining the Burst Buffer concept.

@@

\category{D.4}{Software}{Operating Systems}
\category{D.4.7}{Operating Systems}{Organization and Design}[hierarchical design]

\terms{Design, Performance}

@@

The overall design seeks to offer a byte-granular, multi-version concurrency
control. Through the use of a copy-on-write style mechanism, multiple versions
of a file can be stored in potentially greatly reduced space. By assuming the
client interface will be through an IO library affords a more complicated
interface that offers richer functionality without requiring user code changes.
By managing data in the IOD layer rather than requring writing to centralized
storage will better support the performnce requirements of integrated
application workflows. At a more detailed view, the various layers of the IO
stack each contribute different functionality and performance implications.

The basic architecture incorporates five layers. The top layer is a high level
IO library, such as the demonstration HDF-5 library. The intent is to only have
access to the storage stack through such an API to manage the complexity of
working with the lower layers. Below the user API is an IO forwarding layer
that redirects IO calls from the compute nodes to the IO dispatching layer.
This IO forwarding layer is analogous to the function of the IO nodes in a
BlueGene machine. The next two layers have considerable functionality. The IO
dispatcher (IOD) serves as the primary storage interface for the IO stack and
is the only way to reach the persistent storage array in lower layers.

The core idea for IOD is to provide a way to manage the IO load that is
separate from the compute nodes and the storage array. Communication intensive
activities, such as two-phase, data sieving IO's data rearrangement process,
can be moved to the IOD layer offloading the communication load from the
compute nodes. IOD has three main purposes. First, if the optional burst buffer
is available, it works as a fast cache absorbing write operations for the
slower trickle out to the central storage array. It can also be used to
retrieve objects from the central storage array for more efficient read
operations and offer data filtering to make client reads more efficient.
Second, it offers the transaction mechanism for controlling data set visibility
and to manage faults that would prevent a data set from being used. Third, data
processing operations can be placed in the IOD. These operations are intended
to offer data rearrangement, filtering, and similar operations prior to it
reaching the central storage array.

While these ideas are not necessarily new, they are new twists on best of class
efforts for these techologies. For example, offloading the collective two-phase
data sieving from the compute nodes to reorganize data has proven effective at
reducing the total time for writing data due to fewer participants involved in
the communication patterns~\cite{lofstead:2011:nessie-staging}.  Beyond these
broad items, there are many important details some of which are examined in
more detail below.

The bottom two layers are the Data Access Object Storage (DAOS) and Versioning
Object Storage Device (VOSD). DAOS is the typical interface that represents
the individual objects stored including the semantics related to ``files''.
VSOD is the actual storage mechanism for the objects defined at the DAOS layer.

@@

Along with the analysis of the published design documents, a discussion of the
design philosophy representing the overall intent is presented. This informtion
reprents information that may or moy not have been written down, but is the 
intent of ultimate product. These insights were gained based on personal
converations with the team members discussing some of the potential challenges
with the design as written. These ideas are presented to give a fuller picture
of where the project is going rather than dwelling on the limitations of the
published documents.

@@

The idea of burst buffers were initially explored in the context of data
staging~\cite{abbasi:2007:datatap,Abbasi:2009:datatap,nisar:2008:staging,zheng:2010:predata}.
These initial designs all use extra compute nodes to represent the data storage
buffer given the lack of any dedicated hardware support for this functionality.
The desired outcome of these initial studies is to motivate how such
functionality might be incorporated and the potential benefits.  Later, these
concepts were proposed to be incorporated as part of the IO
stack~\cite{bent:2012:challenges,bent:2012:burst-buffer}.  The current
FastForward IOD design recommends incorporating SSDs, but specifically lists
these devices as optional. Unfortunately, both incorporating burst buffers and
the use of SSDs in the IOD layer may be problematic.  First, with these burst
buffers being optional, the semantics must be clarified of how IOD works must
change to use DAOS to store the data directly rather than storing them locally
until explicitly persisted by the user. The design of DAOS does not incorporate
a high transaction rate mechanism. Conversely, it assumes that it will only be
involved when persisting a completed transaction and only for a fraction of the
total transactions created. Further, with function shipping offering the
ability to change how the data is stored and arranged prior to it being written
to DAOS, this functionality may also be optional depending on the existence of
the burst buffer or slow the IO path to accommodate the data processing.
Consider the important functionality of data rearrangement and doing things
like changing the fast array dimension.

One of the bigger concerns is the observation that the original data staging
proposals all used compute nodes while the newer proposals seek not only to
make them a fixed portion of the IO stack, but also shared across all machine
users. The PreDatA~\cite{zheng:2010:predata} paper in particular examines the
potential costs and advantages of where to place operators similar to the IOD
proposed function shipping. There are two key takeaways from that work. First,
placement matters.  Depending on the communication intensity vs. computation
intensity, where along the IO path to place the operation can matter
significantly. Second, and more importantly, the amount of time spent
processing for the operators was stretched to the point where it consumed
nearly all of the time between IO operations. The given ratios of compute
processes to staging process examined is representative for future extreme
scale platforms. If anything, the ratios offer more staging processes than IOD
processes would be available.

In the case of the written IOD design, it forces a staging area deployment and
then shares the staging area across all users simultaneously. This is unlikely
to be useful because of the limited compute and communication capacity to spare
to perform these operations at a bottleneck in the IO path. The use of a
separate data staging area intentionally separate from the IO path allows using
operators on limited resources leaving the IO path clear for strictly data
movement.

@@

\subsection{Design Philosophy}
The burst buffers design, as presented in the IOD documents, limits the
placement of the function operators and SSDs for buffers to the IO nodes. The
team does acknowledge the limitations of this design and intend to ultimately
focus on spreading the IOD from the IO nodes into the compute area as well.
This is intended both to help address the limitations of the IO bandwidth
and compute capaibility of these few nodes for data processing, but also to
take advantage of new layers in the storage hierarchy. By incorporating NVRAM
into compute nodes, new options for buffering data prior to being moved to
centralized storage become available. This lessens the impact of some operators
while offering additional options for places to store data.

@@

Another concern that is acknowledged, but no thought as been applied to is the
requirement that a single IOD process be the master for any operation. Should
the number of concurrent applications exceed the available nodes, sharing an
IOD process will be required. Since partitioning of the IOD processes for
exclusive use by particular applications is the proposed operating mode, should
insufficient IOD resources be available, either a job could be delayed or IOD
resources could be reallocated from a different process could be redeployed for
use by the new job. These sorts of consideraitons still need to be made for a
full production system.

@@

\subsubsection{Design Philosophy}
Undocumented, but inherent in the design of these transactions is how faults
are detected. The initial design assumes that the current Lustre fault
detection mechanism that can determine if a process or node is no longer
reachable is used. This defines how a fault will be detected and what will trigger a passive fault recovery (i.e., transaction abort).

@@

\subsection{Metadata Management}
Eliminating metaata management as a special case and instead treating it just
as data is a central design goal of the Fast Forward project. Unfortunately,
it is not completely achieved.

@@

\subsubsection{Design Philosophy}
While the metadata design is not fully defined, there are a few things that
are intended as part of this. For example, there is a standard, well known
container that is the system metadata. This includes the list of all other
containers. This container is treated like any other data in the system and
striped as appropriate at both the IOD and the DAOS layers. Unfortunately, this
still couples the metadata to a single object that must serialize access. If
the metadata, including information about striping and other data layout
operations were separated completely from the data path, more scalable
throughput could be achieved. The real challenge of this is introduced by the
the IOD, DAOS, and VOSD layers. Each of these requires some different metadata
storage and the migration is transparent to the user. Supporting fully
independent metadata with this model is difficult, if not impossible. Serious
thought on how to do this outside the data path effectively should be
considered for phase two.

The \DDT project created a simplistic model for a metadata service that has
different features. It is described below and has been published
previously~\cite{lofstead:2012:txn-metadata}. Some of the major differences
include the following. First, \DDTns's metadata service has a way to get a list
of the objects in the metadata service. Since this was intended as a test
system to examine \DDT performance, we did not put the effort into a making it
a robust system. While it does not address having a file equivalence like the
container concept, it does introduce a way to figure out what is available.
Stored along with this object description is a list of the global dimensions
for array objects. One observation made as part of creating this service is that
this is insufficient for many engineering codes that use non-regular meshes.
Second, for each object, a list of the equivalent to shards are kept with a
link back to the master object and the offset and size for each dimension for
that shard. This approach is not scalable because of the explosion in the
process count. Using a striping approach like IOD does is superior. Third,
there is no mapping to another layer with different object layout and counts.

Based on the lessons from the \DDT metadata service construction and the prior
experiments with LWFS, having a completely separate metadata service is
workable. Rather than making it a bottleneck in the IO path, it is another
service that users must interact with if they need those services. Otherwise,
it is 100\% out of the way. Users can manage everything by maintaining the
metadata including the list of objects themselves. However, there are drawbacks
to this 100\% client-side approach.

With a client-side only approach, there is a serious risk of the metadata
service and the object store getting out of sync.  While having a metadata-less
object storage service is desirable, the different semantics from traditional
file systems requires some considerations. In this case, should these services
get out of sync, three particular risks introduced.  First, a client could
create an entry in the metadata service that does not correspond to any objects
in the object store. Second, a client could create objects that have no
associated metadata entries. Third, updates to the metadata or object store
service should be an atomic operation, but due to a lack of coordination, a
window where the system is inconsistent appears.  Ultimately, the consistency
semantics required must be determined. If a metadata service is required and it
must be in sync with the object storage service, then additional work must be
performed. In traditional file systems, the metadata and object storage updates
are atomic. With decoupling metadata from object storage, should this atomicity
still be desired, it requires both the ability for the services to participate
in a task that is part of a larger atomic operation and a higher-level
mechanism to manage the atomic operation. Overall, while additional work is
required to maintain a client-side only metadata service, eliminating any
potential bottlenecks related to updating metadata related to the object
storage are eliminated. The burden of tracking striping and other metadata that
has traditionally been part of the metadata associated with the file system
will have to be maintained by the object storage service. The lack of a
centralized, serialized bottleneck to store that information improves
concurrency.

@@

The example metadata and data storage services created as part of the \DDT
project lack one of the key transaction management features used by IOD.
The major difference is that there are no dependencies between transactions
that prevent visibility should an older version be incomplete. This additional,
intentional requirement by IOD offers different functionality than \DDTns's
example services. In the case of \DDTns, the functionality is more minimal, but
also avoids some of the concerns outlined below.

The second iteration of the protocol~\cite{lofstead:2013:pdsw-txn} fixed
scalability issues and demonstrated a scalable client-side coordination model
with excellent performance. The performance measured for a complex transaction
with \DDT is illustrated in Figure~\ref{fig:performance}. This performance is
explored in detail in a previous paper~\cite{lofstead:2013:pdsw-txn}. The
breakdown of the number of participants in each role is shown in
Table~\ref{tab:scaling}. For comparison, consider the Number of
Sub-Coordinators equivalent to an IOD process. The Processes Per
Sub-Coordinator represents the number of clients that use a particular IOD. For
these tests, we maintained a balanced distribution and always used at least two
sub-coordinators to slow down the processing.

\begin{figure}[ht]
%\vspace{-0.15in}
\centering
\includegraphics[keepaspectratio=true, width=0.9\columnwidth]{images/performance}
\vspace{-0.15in}
\caption{Total Transaction Overhead}
\label{fig:performance}
\vspace{-0.15in}
\end{figure}

@@

At a high level, both \DDT and the IOD transactions have the same high-level
design. In both cases, a hierarchical model is employed. In the case of \DDTns,
it is a purely client-side tree using semi-synchronous messaging. The messaging
itself, in the current implementation, uses asynchronous MPI messages. The
synchronous component comes from the timeout mechanism used to detect faults.
It forces a level of coordination and synchronization for the protocol. For
IOD, it is a server-side tree and fully asynchronous. In both cases, there is a
master in charge of managing the transaction and a collection of workers that
aggregate into the master through second-level leaders. Beyond that, there are
some significant differences. Some of the different choices made by IOD raise
some possible concern.

First, should a transaction on a container not complete, all subsequent
transactions on that container, even if they are complete and correct, will not
be accessible. Once the transaction fails, all of these subsequent transactions
will be aborted. While this is being reconsidered for phase two, it does
complicate the interface to offer the additional transaction mode.

Second, in the multi-leader model, using a count of client connections to
determine if transaction is complete is problematic.  First, should this be in
an environment where lost or repeated network messages are possible, then the
count could be wrong inappropriately. This is a general problem and unlikely
in the current environment, but is a concern on less integrated platforms.
The idea of a count precludes any tracking of expected messages from sources
introducing a potential hole. Second, the aggregation of completion messages
are based on the local aggregation point reporting to the transaction
leader the state.  There is no ability to detect or recover from a failure of
this IOD process that would report to the transaction leader. Third, because a
simple count of participating processes is sent with the beginning of the
transaction, each process is strictly limited to a single output operation.
This precludes a process writing multiple array shards, for example. To put
this in concrete terms, it means that a file could only contain a single,
globally distributed variable (at the root level of HDF5) OR an attribute per
process. If this count actually represents a begin/end transaction pair
instead, it suffers from a lack of accountability for missing object (shard)
writes. All it would indicate is that a process started and ended a transaction
connection. There is no validation that a process has done everything required
of it.

\begin{figure}[ht]
\centering
\includegraphics[keepaspectratio=true, width=0.9\columnwidth]{images/optimized-protocol}
\vspace{-0.15in}
\caption{Optimized Protocol}
\label{fig:optimized-protocol}
\vspace{-0.25in}
\end{figure}

\DDT has addressed this count issue in a couple of ways. First, the
sub-coordinators each have a list of processes from which they expect messages.
Should a message be missed, it is noticed and corrective action can be taken.
Second, \DDT has the concept of sub-transactions. The messaging requirements
are illustrated in Figure~\ref{fig:optimized-protocol}.  Sub-transactions
represent finer grained operations than the entire output, \DDT can manage
multiple writes per client by using a sub-transaction to represent the output
for any item to the file (container). Because of how the sub-transactions are
managed, the singleton sub-transactions, ones in which only a single process
participates, must be declared before the transaction begins so that its
existence can be broadcast as part of the begin transaction message. This
ensures there is global knowledge that the sub-transaction is expected.  That
way if the coordinator (transaction leader) fails, which ever process takes
over that role knows to expect a completion message for that sub-transaction or
the overall transaction cannot complete. Global sub-transactions can be defined
at any time since they are a global, synchronized operation broadcasting their
existence. While this additional layer does introduce messaging, the overhead
is quite small.

@@

\subsection{Comparison to Other Protocols}
Alternatives, such as Paxos~\cite{Lamport:1998:paxos} algorithms like
ZooKeeper~\cite{Hunt:2010:zookeeper}, suffer from two limitations making them
unsuitable for this environment. First, the distributed servers in Paxos
systems are all distributed copies of each other that eventually become
consistent. Given the scale we wish to address, a single node's memory is
unlikely to be able to hold all of the data necessary for many operations at
scale. They also do not have a guarantee for when consensus will be achieved
without using slower synchronous calls. For the tight timing we wish to
support, we need guarantees of when a consistent state has been achieved.
Second, these systems also all assume that updates are initiated from a single
client process rather than a parallel set of processes as is the standard in
HPC environments.

\DDT uses a second layer of coordination on the client side that greatly
increases the scalability by consolidating messages from clients into unique
sets prior to sending to the overall coordinator. A gossip
protocol~\cite{ganesh:2003:gossip-protocols} may appear sufficient for this
purpose, but the delay of eventual consistency is strictly avoided with this
protocol to ensure guarantees at particular states in the code. For example, if
a failure occurs, the global understanding of the role of all processes is
required in order for effective communication to occur for operations like
creating sub-transactions or voting. In this case, the protocol can offer
stronger statements about consistency than these protocols offer.  These
features offer a way to easily scale the transaction protocol given the
guarantees we wish to offer. The IOD approach does nothing to address these
concerns.

Another effort to offer consistency and data integrity for the ZFS file
system~\cite{zhang:2010:zfs} covers some of the same territory. Instead of a
focus on the processes all having a notion of completion as a transaction, this
work focuses on the integrity of the data movement operations. We view this
work as something that should be considered hand-in-hand with a transaction
approach to ensure the integrity of the movement of the data in addition to the
agreement of processes about the successful completion of a parallel operation.

>>>>

\documentclass[conference]{IEEEtran}
\pdfpagewidth=8.5in
\pdfpageheight=11in
%\usepackage{subfig}
\usepackage{subfigure}
\usepackage[pdftex]{}
\usepackage{todonotes}
\usepackage{listings}
\lstset {% general command to set parameter(s)
         language=C,
     basicstyle=\footnotesize,               % print whole listing small
%     keywordstyle=\color{black}\bfseries, % underlined bold black keywords
%     identifierstyle =\color{black},  % nothing happens
%     commentstyle=\color{black}\emph, % white comments
     %stringstyle=\ttfamily,          % typewriter type for strings
%     stringstyle=\color{black},       % typewriter type for strings
         tabsize=4,
         showtabs=false,
     showstringspaces=false}
%can't figure this one out for particles bandwidth
%\usepackage[caption,label]{subfig}

@@

\addtolength{\parskip}{-0.04in}

@@

%\conferenceinfo{MSST'14,} {November 17, 2013, Denver, Colorado, USA.}
%\CopyrightYear{2014}
%\crdata{978-1-4503-1103-8/11/11}
%\clubpenalty=10000
%\widowpenalty = 10000

@@

\author{
\IEEEauthorblockN{Jay Lofstead}
\IEEEauthorblockA{Sandia National Laboratories\\gflofst@sandia.gov}
\and
\IEEEauthorblockN{Ivo Jimenez}
\IEEEauthorblockA{University of California, Santa Cruz\\ivo@cs.ucsc.edu}
\and
\IEEEauthorblockN{Carlos Maltzahn}
\IEEEauthorblockA{University of California, Santa Cruz\\carlosm@soe.ucsc.edu}
}
\maketitle

@@

While these and other their contributions are valuable, similar efforts made
elsewhere may offer attractive alternatives or differing semantics that could
yield a more feature rich environment with little to no additional overhead.
For example, the Doubly Distributed Transactions (\DDTns) protocol offers an
alternative approach for incorporating transactional semantics into the data
path. The PreDatA project examined how to get the best throughput for data
operators can offer some insights for refining the Burst Buffer concept.

@@

%\category{D.4}{Software}{Operating Systems}
%\category{D.4.7}{Operating Systems}{Organization and Design}[hierarchical design]

%\terms{Design, Performance}

@@

The overall design seeks to offer a byte-granular, multi-version concurrency
control. Through the use of a copy-on-write style mechanism, multiple versions
of a file can be stored in potentially greatly reduced space. It assumes the
client interface will be through an IO library affording a more complicated
interface that offers richer functionality without requiring user code changes.
Managing most data access in a machine-local layer rather than requiring
writing to centralized storage will better support the performance and energy
requirements of extreme scale integrated application workflows. At a more
detailed view, the various layers of the IO stack each contribute different
functionality and performance implications.

The basic architecture incorporates five layers. The top layer is a high level
IO library, such as the demonstration HDF-5 library. The intent is to only have
access to the storage stack through such an API to manage the complexity of
working with the lower layers. Below the user API is an IO forwarding layer
that redirects IO calls from the compute nodes to the IO dispatching layer.
This IO forwarding layer is analogous to the function of the IO nodes in a
BlueGene machine. The next two layers have considerable functionality. The IO
dispatcher (IOD) serves as the primary storage interface for the IO stack and
is the only way to reach the persistent storage array in lower layers. The
Data Access Object Storage (DAOS) layer serves as the persistent storage
interface. At the bottom is the Versioning Object Storage Device (VOSD). It
serves as the interface for storing objects of all types efficiently.

@@

The core idea for IOD is to provide a way to manage the IO load that is
separate from the compute nodes and the storage array. Communication intensive
activities, such as two-phase, data sieving IO's data rearrangement process,
can be moved to the IOD layer offloading the communication load from the
compute nodes. IOD has three main purposes. First, if the optional burst buffer
is available, it works as a fast cache absorbing write operations for the
slower trickle out to the central storage array. It can also be used to
retrieve objects from the central storage array for more efficient read
operations and offer data filtering to make client reads more efficient.
Second, it offers the transaction mechanism for controlling data set visibility
and to manage faults that would prevent a data set from being used. Third, data
processing operations can be placed in the IOD. These operations are intended
to offer data rearrangement, filtering, and similar operations prior to data
reaching the central storage array.

While these ideas are not necessarily new, they are new twists on best of class
efforts for these technologies. For example, offloading the collective
two-phase data sieving from the compute nodes to reorganize data has proven
effective at reducing the total time for writing data due to fewer participants
involved in the communication patterns~\cite{lofstead:2011:nessie-staging}.
Beyond these broad items, there are many important details some of which are
examined in more detail below.

Along with the analysis of the published design documents, a discussion of the
design philosophy representing the overall intent is presented. This
information represents information that may or may not have been written down,
but is the intent of ultimate product. These insights were gained based on
personal conversations with the team members discussing some of the potential
challenges with the design as written. These ideas are presented to give a
fuller picture of where the project is going rather than dwelling on the
limitations of the published documents.

@@

The idea of burst buffers were initially explored in the context of data
staging~\cite{abbasi:2007:datatap,Abbasi:2009:datatap,nisar:2008:staging,zheng:2010:predata}.
These initial designs all use extra compute nodes to represent the data storage
buffer given the lack of any dedicated hardware support for this functionality.
The desired outcome of these initial studies is to motivate how such
functionality might be incorporated and the potential benefits.  Later, these
concepts were proposed to be incorporated as part of the IO
stack~\cite{bent:2012:challenges,bent:2012:burst-buffer}.  The current
FastForward IOD design recommends incorporating SSDs, but specifically lists
these devices as optional. Unfortunately, both incorporating burst buffers and
the use of SSDs in the IOD layer may be problematic.  First, the IOD and DAOS
designs currently are written assuming burst buffers. This means that the
bulk of IO operations will only hit the IOD layer and proposed functionality,
such as the function operators, do not discuss the impact on the DAOS layer or
the operators themselves should a burst buffer not be available.  Consider the
important functionality of data rearrangement and doing things like changing
the fast array dimension on shared, spinning media.  The design of DAOS assumes
that it will only be involved when persisting a completed transaction and only
for a fraction of the total transactions created.

One of the bigger concerns is the observation that the original data staging
proposals all used compute nodes while the newer proposals seek not only to
make them a fixed portion of the IO stack, but also shared across all machine
users. The PreDatA~\cite{zheng:2010:predata} paper in particular examines the
potential costs and advantages of where to place operators similar to the IOD
proposed function shipping. There are two key takeaways from PreDatA. First,
placement matters.  Depending on the communication intensity vs. computation
intensity, where along the IO path to place the operation can matter
significantly. Second, and more importantly, the amount of time spent
processing for the operators was stretched to the point where it consumed
nearly all of the time between IO operations. The given ratios of compute
processes to staging process examined is representative for future extreme
scale platforms. If anything, the ratios offer more staging processes than IOD
processes would be available.

In the case of the written IOD design, it describes a fixed-sized staging area
shared across all users simultaneously.  This is unlikely to be useful because
of the limited compute and communication capacity to spare to perform these
operations at a bottleneck in the IO path.  The use of a separate data staging
area intentionally separate from the IO path allows using operators on limited
resources leaving the IO path clear for strictly data movement. A nuance of
this design is discussed in the Design Philosophy below.

@@

\subsection{Design Philosophy}
The burst buffers design, as presented in the IOD documents, limits the
placement of the function operators and SSDs for buffers to the IO nodes. The
team does acknowledge the limitations of this design and intend to ultimately
focus on spreading the IOD layer from the IO nodes into the compute area as
well.  This is intended both to help address the limitations of the IO
bandwidth and compute capability of these few nodes for data processing, but
also to take advantage of new layers in the storage hierarchy. By incorporating
NVRAM into compute nodes, new options for buffering data prior to being moved
to centralized storage become available and addresses some of the concerns
about SSD performance. This lessens the impact of some operators while offering
additional options for places to store data.

@@

Another concern that is acknowledged, but no thought as been applied to is the
requirement that a single IOD process be the master for any operation. Should
the number of concurrent applications exceed the available nodes, sharing an
IOD process will be required. Since partitioning of the IOD processes for
exclusive use by particular applications is the assumed operating mode, should
insufficient IOD resources be available, either a job could be delayed or IOD
resources could be reallocated from a different process could be redeployed for
use by the new job. These sorts of considerations still need to be made for a
full production system.

@@

\subsubsection{Design Philosophy}
Undocumented, but inherent in the design of these transactions is how faults
are detected. The initial design assumes the current Lustre fault detection
mechanism that can determine if a process or node is no longer reachable. This
defines how a fault will be detected and what will trigger a passive fault
recovery (i.e., transaction abort).

@@

\subsection{Metadata Management}
Eliminating metadata management as a special case and instead treating it just
as data is a central design goal of the Fast Forward project. Unfortunately,
it is not completely achieved.

@@

\subsubsection{Design Philosophy}
While the metadata design is not fully defined, there are a few things that
are intended. For example, there is a standard, well-known container that is
the system metadata. This includes the list of all other containers. This
container is treated like any other data in the system and striped as
appropriate at both the IOD and the DAOS layers. Unfortunately, this still
couples the metadata to a single object that must serialize access. If the
metadata, including information about striping and other data layout operations
were separated completely from the data path, more scalable throughput could be
achieved. The real challenge of this is introduced by the IOD, DAOS, and VOSD
layers. Each of these requires some different metadata storage and the
migration is transparent to the user. Supporting fully independent metadata
with this model is difficult. Serious thought on how to do this effectively
outside the data path should be considered for phase two.

Based on the lessons from the \DDT metadata
service~\cite{lofstead:2012:txn-metadata} construction and the prior
experiments with LWFS, having a completely separate metadata service is
workable. Rather than making it a bottleneck in the IO path, it is another
service that users must interact with if they need those services.  Users can
manage everything by maintaining the metadata including the list of objects
themselves. However, there are drawbacks to this 100\% client-side approach.

With a client-side only approach, there is a serious risk of the metadata
service and the object store getting out of sync.  While having a metadata-less
object storage service is desirable, the different semantics from traditional
file systems requires some considerations. In this case, should these services
get out of sync, three particular risks introduced.  First, a client could
create an entry in the metadata service that does not correspond to any objects
in the object store. Second, a client could create objects that have no
associated metadata entries. Third, updates to the metadata or object store
service should be an atomic operation, but due to a lack of coordination, a
window where the system is inconsistent appears.  Ultimately, the consistency
semantics required must be determined. If a metadata service is required and it
must be in sync with the object storage service, then additional work must be
performed. In traditional file systems, the metadata and object storage updates
are atomic. With decoupling metadata from object storage, should this atomicity
still be desired, it requires both the ability for the services to participate
in a task that is part of a larger atomic operation and a higher-level
mechanism to manage the atomic operation. Overall, while additional work is
required to maintain a client-side only metadata service, it eliminates any
potential bottlenecks related to updating metadata related to the object
storage. The burden of tracking striping and other metadata that has
traditionally been part of the metadata associated with the file system will
have to be maintained by the object storage service. The lack of a centralized,
serialized bottleneck to store that information improves concurrency.

@@

The example metadata and data storage services created as part of the \DDT
project have no dependencies between transactions that prevent visibility
should an older version be incomplete. This additional, intentional requirement
by IOD offers different functionality than \DDTns's example services. In the
case of \DDTns, the functionality is more minimal, but also avoids some of the
concerns outlined below.

The second iteration of the protocol~\cite{lofstead:2013:pdsw-txn} fixed
scalability issues and demonstrated a scalable client-side coordination model
with excellent performance. The performance measured for a complex transaction
with \DDT is illustrated in Figure~\ref{fig:performance}. This performance is
explored in detail in a previous paper~\cite{lofstead:2013:pdsw-txn}. The
breakdown of the number of participants in each role is shown in
Table~\ref{tab:scaling}. For comparison, consider the Number of
Sub-Coordinators equivalent to IOD processes. The Processes Per Sub-Coordinator
represents the number of clients that use a particular IOD process. For these
tests, we maintained a balanced distribution and always used at least two
sub-coordinators to slow down the processing.

\begin{figure}[ht]
%\vspace{-0.15in}
\centering
\includegraphics[keepaspectratio=true, width=0.9\columnwidth]{images/performance}
\vspace{-0.15in}
\caption{Total Transaction Overhead}
\label{fig:performance}
%\vspace{-0.15in}
\end{figure}

@@

At a high level, both \DDT and the IOD transactions have the same design. In
both cases, a hierarchical model is employed. In the case of \DDTns, it is a
purely client-side tree using semi-synchronous messaging. The messaging itself,
in the current implementation, uses asynchronous MPI messages. The synchronous
component comes from the timeout mechanism used to detect faults.  It forces a
level of coordination and synchronization for the protocol. For IOD, it is a
server-side tree and fully asynchronous relying on the existing Lustre fault
detection mechanism for failure detection. In both cases, there is a master in
charge of managing the transaction and a collection of workers that aggregate
into the master through second-level leaders. Beyond that, there are some
significant differences. Some of the different choices made by IOD raise some
possible concern.

In the multi-leader model, using a count of client connections to determine if
transaction is complete is problematic.  First, the aggregation of completion
messages are based on the local aggregation point reporting to the transaction
leader.  There is no ability to detect or recover from a failure of this IOD
process that would report to the transaction leader. The current assumption
that the IOD layer will be an MPI program is not workable until MPI can fully
support fault detection and recovery including removing and adding processes.
Second, because a simple count of participating processes is sent with the
beginning of the transaction, each process is strictly limited to a single
output operation.  This precludes a process writing multiple array shards, for
example. To put this in concrete terms, it means that a file could only contain
a single, globally distributed variable (at the root level of HDF5) OR an
attribute per process. If this count actually represents a begin/end
transaction pair instead, it suffers from a lack of accountability for missing
object (shard) writes. All it would indicate is that a process started and
ended a transaction connection. There is no validation that a process has done
everything required of it.

\begin{figure}[ht]
\centering
\includegraphics[keepaspectratio=true, width=0.9\columnwidth]{images/optimized-protocol}
\vspace{-0.15in}
\caption{Optimized Protocol}
\label{fig:optimized-protocol}
%\vspace{-0.25in}
\end{figure}

\DDT has addressed theses issues in a couple of ways. First, the
sub-coordinators each have a list of processes from which they expect messages.
Should a message be missed, it is noticed and corrective action can be taken.
Second, \DDT has the concept of sub-transactions. The messaging requirements
are illustrated in Figure~\ref{fig:optimized-protocol}.  Sub-transactions
represent finer grained operations than the entire output, \DDT can manage
multiple writes per client by using a sub-transaction to represent the output
for any item to the file (container). Because of how the sub-transactions are
managed, the singleton sub-transactions, ones in which only a single process
participates, must be declared before the transaction begins so that its
existence can be broadcast as part of the begin transaction message. This
ensures there is global knowledge that the sub-transaction is expected.  That
way if the coordinator (transaction leader) fails, which ever process takes
over that role knows to expect a completion message for that sub-transaction or
the overall transaction cannot complete. Global sub-transactions can be defined
at any time since they are a global, synchronized operation broadcasting their
existence. While this additional layer does introduce messaging, the overhead
is quite small.

@@

%\subsection{Comparison to Other Protocols}
%Alternatives, such as Paxos~\cite{Lamport:1998:paxos} algorithms like
%ZooKeeper~\cite{Hunt:2010:zookeeper}, suffer from two limitations making them
%unsuitable for this environment. First, the distributed servers in Paxos
%systems are all distributed copies of each other that eventually become
%consistent. Given the scale we wish to address, a single node's memory is
%unlikely to be able to hold all of the data necessary for many operations at
%scale. They also do not have a guarantee for when consensus will be achieved
%without using slower synchronous calls. For the tight timing we wish to
%support, we need guarantees of when a consistent state has been achieved.
%Second, these systems also all assume that updates are initiated from a single
%client process rather than a parallel set of processes as is the standard in
%HPC environments.
%
%\DDT uses a second layer of coordination on the client side that greatly
%increases the scalability by consolidating messages from clients into unique
%sets prior to sending to the overall coordinator. A gossip
%protocol~\cite{ganesh:2003:gossip-protocols} may appear sufficient for this
%purpose, but the delay of eventual consistency is strictly avoided with this
%protocol to ensure guarantees at particular states in the code. For example, if
%a failure occurs, the global understanding of the role of all processes is
%required in order for effective communication to occur for operations like
%creating sub-transactions or voting. In this case, the protocol can offer
%stronger statements about consistency than these protocols offer.  These
%features offer a way to easily scale the transaction protocol given the
%guarantees we wish to offer. The IOD approach does nothing to address these
%concerns.
%
%Another effort to offer consistency and data integrity for the ZFS file
%system~\cite{zhang:2010:zfs} covers some of the same territory. Instead of a
%focus on the processes all having a notion of completion as a transaction, this
%work focuses on the integrity of the data movement operations. We view this
%work as something that should be considered hand-in-hand with a transaction
%approach to ensure the integrity of the movement of the data in addition to the
%agreement of processes about the successful completion of a parallel operation.
